---
title: "P8130_final_project"
author: "Yifei Yu"
date: "2024-12-03"
output: pdf_document
editor_options: 
  chunk_output_type: console
  fontsize: 12pt
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract


## Introduction

The data we used for this analysis originates from a breast cancer survival dataset collected from a prospective study. The dataset contains 14 important variables, including patients’ age, race, marital status, tumor size, cancer stages (T Stage, N Stage, and 6th Stage), estrogen and progesterone status, and regional node involvement. The dataset also records patients’ survival times in months and their final survival status (dead or alive). Our objective is to develop models that predict the risk of death among breast cancer patients using these features. Specifically, we aim to identify which variables significantly impact patients’ survival outcomes, determine (potential interactions among the variables, and assess the performance of the models across racial groups). Additionally, we will investigate fairness in the model predictions to ensure equitable accuracy between the majority race group “White” and the minority “Black”.

## Methods


This report was conducted using `survival_data`, which contains information about breast cancer patients from a prospective study. The dimension of this dataset is `r nrow(survival_df)` rows * `r ncol(survival_df)` columns. It contains five numeric variables, which are `age`, `tumor_size`, `regional_node_examined`, `reginol_node_positive`, and `survival_months`, and eleven categorical variables, which are `race`, `marital_status`, `t_stage`, `n_stage`, `x6th_stage`, `differentiate`, `grade`, `a_stage`, `estrogen_status`, `progesterone_status`, and `status`. Key variables in the dataset include:

**age**: The age of the patient (in years).

**differentiate**: Tumor differentiation grade, categorized as "Well differentiated", "Moderately differentiated", "Poorly differentiated", or "Undifferentiated".

**a_stage**: Categorized as "Regional" (a neoplasm that has extended) or "Distant" (a neoplasm that has spread to
parts of the body remote from).

**tumor_size**: The size of tumor (in millimeters).

**regional_node_examined**: The number of examined regional nodes.

**regional_node_positive**: The number of positive reginol nodes.

**survival_month**: The time of a patient with breast cancer is expected to live after their diagnosis (in months).


The dataset comprises a total of `r count(survival_df)` observations, with no missing values in the primary variables analyzed. Firstly, after cleaning and tidying the data, we could figure out the distributions of the data and check for potential outliers or influential points by plotting the distributions of the variables, such as histograms and box plots. We also examine the pairwise relationships between variables.




## Results


## Conclusion/Discussion


## Group members' contribution

Leyang and Jinghan focused on statistical methods, building models to extract meaningful insights. Yuechu and Yifei concentrated on data description and visualization, presenting information through clear visuals. All four of us contributed to the writing of the report, integrating our individual efforts into a cohesive final report that reflects both analytical depth and visual clarity.



## Appendix



```{r, echo=FALSE, fig.cap="Distribution of the Continuous Variables", include=FALSE}
survival_df |> 
  pivot_longer(
    cols = c(age, tumor_size, regional_node_examined, reginol_node_positive, survival_months),
    names_to = "variable",
    values_to = "value"
  ) |>
  ggplot(aes(x = value)) +
  geom_histogram(color = "light blue") +
  facet_wrap(variable ~ .,  scales = "free") +
    labs(title = "Distribution of the Continuous Variables",
         x = "Variables",
         y = "Frequency")
```






